---
title: "Dust loading on Colorado Snowpack"
author: "Billy Johnson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Intro
### Load in Packages
```{r}
library(tidyverse)
library(kableExtra)
library(ggthemes)
library(readxl)
library(readr)
library(purrr)
library(magrittr)
library(lubridate)
library(scales)
library(viridis)
library(pals)
library(plotly)
library(ggplot2)
library(patchwork)
library(broom)

knitr::opts_chunk$set(echo = TRUE)
theme_set(theme_few())
```

### Objectives and Research Question

How does dust loading in the Colorado snow pack affect alpine areas energy budget?

•	Examine how dust events influence outgoing short and long-wave radiation, as well as surface snow temperature.

•	Evaluate melt rates when dust layers are exposed versus not exposed


## Loading in the Data --------------------------
```{r}
# Load in Dust Mass Data
SASP_Mass_Simple <- read_excel("Data/In/SASP_Mass_Simple.xlsx")

#Load in the Energy Budget Data 24 hrs
Energy_24hr <- read_csv("Data/In/SASP_24hr.csv")

# Load in the energy budget data for 1 hr
Energy_1hr <- read_csv("Data/In/SASP_1hr_2010-2023.csv")


# Load in the 24 hr Stream gauge data
stream_24hr <- read_csv("Data/In/SBSG_24hr.csv")

# Load in the 1 hr stream gauge data
stream_1hr <- read_csv("Data/In/SBSG_1hr (1).csv")

# Load in SWE
SWE <- read_csv("Data/In/SWE.txt",
                col_names = c("Date", "SWE_in", "Precip_in", "Air_Temp_Max_F", "Air_Temp_Min_F", "Air_Temp_Avg_F", "Precip_Incriment_in"),
                   skip = 64)
```

# Cleaning
## Dust Data
```{r}
# Get sheet names
dust_sheets <- excel_sheets("Data/In/SASP_Mass_Simple.xlsx") 


# SASP_Mass_Loading_Data <- read_excel("Data/In/SASP+Mass+Loading+Data.xls")
dust_data_years <- dust_sheets %>%
  # Split the incoming strings on a space to grab years
  str_split_fixed(., ' ', 3) %>%
  # grab only the year column
  .[,2] %>%
  #convert to numeric
  as.numeric(.) 


# Combine the numeric year column and the sheet name column
sheet_table <- tibble(year = dust_data_years,
                      sheet = dust_sheets) %>%
  filter(year != 2021)

sheet_table

# Getting all sheets into R         ? I am having issue becase I removed 2021 from the data and now the if else function is not working right, i belive that I just have a typing issue with the brackets

# Creating a function to get a sheet into R.
custom_reader <- function(sheet = sheet_table$sheet[1], year = sheet_table$year[1]){
  read_excel('Data/In/SASP_Mass_Simple.xlsx',
             # The first ^ rows just have text we want to skip those
             sheet = sheet, skip = 6,
             col_types = 'text') %>%
    mutate(year = year)
}

# # Create an Empty list for the loop function
# empty_list <- list()
# 
# # Loop function to get all sheets
# for(i in 1:nrow(sheet_table)){
#   # Creating values
#   empty_list[[i]] <- custom_reader(sheet = sheet_table$sheet[i],
#                                    year = sheet_table$year[i])
#   print(i)
# }
# 
# 
# # Stack the list of sheets into one table
# dust_analysis <- do.call('rbind',empty_list)
# 

dust_2 <- pmap_dfr(sheet_table, custom_reader)

dust_mass <- dust_2 %>% 
  filter(grepl("^\\d+\\.?\\d*$", Mass))

dust_mass %<>% 
  mutate(date = lubridate::mdy(DateClean))

dust_mass %>% 
  mutate(DateClean = as.Date(DateClean, format = "'%m/%d/%Y'"))

unique(dust_2$Weigth)
View(dust_2)
str(dust_2)

write_csv(dust_mass, "Data/Out/Dust_Clean.csv")
Dust_clean <- read_csv("Data/Out/Dust_Clean.csv")
```

### Changing the column types for dust data
```{r}
dust_2 %<>% 
  mutate(date = lubridate::mdy(DateClean))

dust_2 %<>%
  mutate(weight = ifelse(is.na(Weigth),NA, as.numeric(Weigth)))

```

## Stream Data 24 hr
```{r}
view(stream_24hr)


stream_24hr_clean <- stream_24hr %>% 
  filter(!(Year >= "2005" & Year <= "2010"))
  
stream_24hr_clean <- stream_24hr %>% 
  filter(!(Year >= "2020" & Year <= "2023"))

stream_24hr_clean <- stream_24hr_clean %>% 
    mutate(Hour = sub("00$", "", Hour),
         date = as.POSIXct(paste(Year, DOY, Hour), format = "%Y %j %H"),
         month = format(date, "%m"),
         day = format(date, "%d"),
         year = format(date, "%Y"))

  
stream_max <- stream_24hr %>% 
  group_by(Year) %>% 
  summarise(max_q = max(Discharge_CFS, na.rm = TRUE),
            total_q = sum(Discharge_CFS, na.rm = TRUE),
            avg_q = mean(Discharge_CFS, na.rm = TRUE))

stream_max <- stream_max %>% 
  mutate(mm = max_q * 2446575580800)
  
  
```

## Stream Data 1 hr
```{r}
stream_1hr_clean <- stream_1hr %>% 
  filter(!(Year >= "2005" & Year <= "2010")) %>% 
  filter(!(Year >="2020" & Year <= "2023"))

stream_1hr_clean <- stream_1hr_clean %>% 
  mutate(Hour = sub("00$", "", Hour),
         date = as.POSIXct(paste(Year, DOY, Hour), format = "%Y %j %H"),
         month = format(date, "%m"),
         day = format(date, "%d"),
         year = format(date, "%Y"))

view(stream_1hr_clean)

```

## Energy Data 
### 1 hr data
```{r}
Energy_1hr_clean <- Energy_1hr %>%
  filter(Year < 2020) %>%
  filter(Year > 2010) %>% 
  mutate(Hour = sub("00$", "", Hour),
         date = as.POSIXct(paste(Year, DOY, Hour), format = "%Y %j %H"),
         month = format(date, "%m"),
         day = format(date, "%d"),
         year = format(date, "%Y"))


glimpse(Energy_1hr_clean)
view(Energy_1hr_clean)
str(Energy_1hr_clean)

write_csv(Energy_1hr_clean, "Data/Out/Energy_1hr_clean.csv")

Energy_1hr_clean <- read_csv("Data/Out/Energy_1hr_clean.csv")
```


#### Calculating Albedo
```{r}
view(Energy_1hr_clean)

Energy_1hr_clean$albedo <-  Energy_1hr_clean$PyDwn_Filt_W/Energy_1hr_clean$PyUp_Filt_W


#albedo_data <- Energy_1hr_clean %>% 
#mutate(albedo = PyDwn_Unfilt_W/)
```

### 24 hr Data
```{r}
View(Energy_24hr)

# View the variables
SASP_varibale_names <- Energy_24hr %>% 
  names()

view(SASP_varibale_names)

# Clean the Energy Budget Data

Energy_clean <- Energy_24hr %>% 
  filter(!(Year >= "2003" & Year <= "2010"))

Energy_clean <- Energy_clean %>% 
  filter(!(Year >= "2020" & Year <= "2022"))

##  Make this a factor to make it categorical for the box plot
Energy_clean <- Energy_clean %>% 
  mutate(Year = factor(Year))  

# Remove the non-numeric numbers
Energy_clean <- Energy_clean %>% 
  filter(!is.na(Pyrgeom_W) & is.finite(Pyrgeom_W))

View(Energy_clean)


```

### SWE
```{r}
SWE_max <- SWE %>% 
  group_by(Year = lubridate::year(Date)) %>% 
  summarise(max_swe = max(SWE_in, na.rm = TRUE))



```

### Time to Peak SWE
```{r}
view(SWE)
```

# Analysis

## Average Snowfall
```{r}
average_snow <- SWE %>%
  mutate(month = lubridate::month(Date),
         year = lubridate ::year(Date)) %>% 
  filter(month %in% c (10,11,12,1,2,3,4)) %>% 
           select(-month) %>%  
  group_by(year) %>% 
  summarise(yearly_avg = mean(Precip_in, na.rm = TRUE))


view(average_snow)

max(average_snow$yearly_avg)
```

## Join Max SWE and Discharge
```{r}
# Create year total for dust accumulation
dust_mass <- dust_mass %>% 
  mutate(Mass = as.numeric(Mass), Mass)


year_total <- dust_mass %>% 
  group_by(Year = year) %>% 
  summarise(total_dust = sum(Mass))

max_dates <- dust_mass %>%
    mutate(doy = yday(date)) %>%
    group_by(Year = year) %>%
    mutate(cume_dust = cumsum(Mass),
           maxday = max(cume_dust, na.rm =T)) %>%
    filter(cume_dust == maxday,
           !is.na(doy)) %>%
  select(year,doy)
    


# Join the MAX SWE and Discharge with Dust totals
Max_SWE_DIS <- inner_join(SWE_max, stream_max, by = "Year") %>% 
  left_join(year_total) %>%
  left_join(max_dates)

```
# Figures ---------------------

## SWE and Q
```{r}
Max_SWE_DIS %>% 
  mutate(mm = avg_q * 2446575580800) %>% 
  mutate(cm = max_swe * 2.54) %>% 
  ggplot(aes(x = max_swe, y = mm, color = total_dust))+
  geom_point(size = 3)+
  scale_color_gradient(low = "grey65", high = "blue")+
  geom_smooth( method = "lm")+
  labs(
       y = " Average Runoff (mm/d)",
       x = " Maximum SWE (cm)",
       color = "Dust Accumulation")

max_nona <- Max_SWE_DIS %>%
  na.omit()

mod <- lm(avg_q ~ max_swe + total_dust, data = max_nona)


mod

summary(mod)
print(mod)

tidy_table <- tidy(mod)

kable(tidy_table)
```

### Residuals
```{r}

q_swe_resid <- max_nona %>%
  mutate(pred_q = predict(mod, max_nona),
         resid = avg_q - pred_q)

ggplot(q_swe_resid, aes(x = doy,
                        y = resid)) + 
  geom_point()

summary(lm(resid ~ total_dust, data = q_swe_resid))



max_nona$pred_q <- predict(mod, max_nona) 



Max_SWE_DIS$residuals <- mod$residuals

ggplot(Max_SWE_DIS, aes(total_dust, residuals))+
  geom_point(shape =1)+
  geom_smooth(method = "lm",
              color = "black")+
  labs(x = "Total Dust Accumulation",
       y = "Residuals")
```

## Stream Data 24 hr
```{r}
stream_24hr_clean %>% 
  ggplot(aes(x = DOY, y = Discharge_CFS))+
  geom_line()+
  facet_wrap(stream_24hr_clean$Year)

```

## Stream Data 1 hr
```{r}
view(stream_1hr_clean)

stream_1hr_clean %>% 
  filter(Year == 2017) %>% 
  ggplot(aes(x = date, y = Discharge_CFS))+
  geom_line()
```

## Dust Data
```{r}
view(dust_2)
str(dust_2)

# Create a dataframe for the total dust accumulations per year
dust_mass <- dust_mass %>% 
  mutate(Mass = as.numeric(Mass), Mass)

year_total <- dust_mass %>% 
  group_by(Year = year) %>% 
  summarise(total_dust = sum(Mass))

yearly_average <- year_total %>% 
  summarise(mean(total_dust))

# Create a dateframe with the total dust accumulations per month
month_total <- dust_mass %>% 
  group_by(month = month(date)) %>% 
  summarise(monthly_total = sum(Mass))

year_total %>% 
  # Set the data to each axis
  ggplot(aes(x = Year, y = total_dust))+ 
  # Make it a bar graph and split them by year
  geom_bar(stat = "identity", fill = "#154360")+
  # Add labels to each bar to help with context and adjust the position
  geom_text(aes(label = total_dust), size = 3, vjust = -1)+
  # Add titles
  labs(
       x = "Year",
       y = "Mass (grams)")+
  theme_minimal()+
   # Add label for each year
  theme(axis.text = element_text(angle = 45, colour = "black", size = 10),  
        legend.position = "none",               # Remove the legend
        panel.grid.major = element_blank(),     # Remove the grid lines
        panel.grid.minor = element_blank(),
        #plot.background = element_rect(fill = "light"),
        plot.title = element_text(hjust = 0.5))+ # Move the title to center 
  scale_x_continuous(breaks = year_total$Year,
                     labels = year_total$Year)+
  scale_y_continuous(labels = function(total_dust) format (total_dust, nsamll = 0.1))

# Graph for monthly data
month_total %>% 
  ggplot(aes(x = month, y = monthly_total))+ 
  geom_bar(stat = "identity", fill = "#169294")+
  geom_text(aes(label = monthly_total), size = 3, vjust = -1)+
  labs(
       x = "Month",
       y = "Mass (grams)")+
  theme_minimal()+
  theme(axis.text = element_text(angle = 45, colour = "black", size = 10),
        legend.position = "none",               
        panel.grid.major = element_blank(),     
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))+ 
  scale_x_continuous(breaks = month_total$month,
                     labels = month_total$month)+
  scale_y_continuous(labels = function(monthly_dust) format (monthly_dust, nsamll = 0.1))
```

### Closer Look at 2013
```{r}
dust_2013 <- dust_mass %>% 
  filter(year == 2013) %>% 
  mutate(DateClean = as.Date(DateClean, format = "'%m/%d/%Y'"))

dust_2013 %>% 
  mutate(Weight = as.numeric(Weigth)) %>% 
  ggplot(aes(x = DateClean, y = Weight))+
  geom_point()+
  scale_y_log10()
```



## Energy Budget data 1 hr
### Plotly
```{r}
Energy_1hr_clean %>%
  filter(date > "2015-04-01" & date < "2015-04-30") %>% 
  plot_ly(x = ~date) %>%  
  add_trace(y = ~PyUp_Unfilt_W, type = 'scatter', mode = 'lines', 
            line = list(color = "black"), name = "Shortwave Radiation") %>% 
  add_trace(y = ~Sno_Height_M, type = 'scatter', mode = 'lines',
            line = list(color = "red"), name = "Snow Depth") %>% 
  add_trace(y = ~Sno_IR_C, type = 'scatter', mode = 'lines', 
            yaxis = 'y2', line = list(color = 'blue'), 
            name = "Surface Temperature") %>% 
  layout(yaxis2 = list(overlaying = "y", side = "right"))


my_colors <- c('red', 'blue', 'black','orange')

graph_rad <- Energy_1hr_clean %>% 
  filter(date > "2015-04-01" & date < "2015-04-30") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = PyUp_Unfilt_W, color = "PyUp_Unfilt_W"))+
  geom_line(aes(y = PyDwn_Unfilt_W, color = "PyDwn_Unfilt_W"))+
  scale_color_manual(values = my_colors[1:2],
                     breaks = c("PyUp_Unfilt_W", "PyDwn_Unfilt_W"),
                     labels = c("Downwelling Shortwave", "Upwelling Shortwave"))+
  theme_few()+
  labs( y= "Downwelling Shortwave Flux",
        color = "Parameters")+
  scale_y_continuous(
    sec.axis = sec_axis(~., name = "Upwelling Shortwave Flux")
  )

graph_rad
rescale_temp <- function(x) { x * 0.1}
graph_sno <- Energy_1hr_clean %>% 
  filter(date > "2015-04-01" & date < "2015-04-30") %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = Sno_Height_M, color = "Sno_Height_M"))+
  geom_line(aes(y = Sno_IR_C, color = "Sno_IR_C"))+
  scale_color_manual(values = my_colors[3:4],
                     breaks = c("Sno_Height_M", "Sno_IR_C"),
                     labels = c("Snow Depth (m)", "Surface Temperature (C)"))+
  theme_few()+
  labs( y = "Snow Height (m)",
        color = "Parameters")+
  scale_y_continuous(sec.axis = sec_axis(~rescale_temp(.), name = "Surface Temperature (C)"))

graph_sno
  
stacked_plots <- graph_rad / graph_sno

stacked_plots
```

### April 2015
```{r}
Energy_1hr_clean %>% 
  filter(month(date) == 04 & year(date) == 2015) %>% 
  ggplot(aes(x = date))+
  geom_line(aes(y = PyUp_Unfilt_W/100), color = "black")+
  geom_line(aes(y = Sno_Height_M), color = "red")+
  geom_line(aes(y = Sno_IR_C), color = "orange")+
  scale_y_continuous(name = "Shortwave Radiation")+
  labs(title = "April 2015")


April_dust <- Dust_clean %>% 
  filter(month(date) == 04 & year(date) == 2015)

April_energy <- Energy_1hr_clean %>% 
  filter(month(date)== 04 & year(date) == 2015)


April_dust %>% 
  ggplot(aes(x = date, y = Mass))+
  geom_point()+
  geom_label(aes(label = date), vjust = 1.5, hjust = 0.5, size = 3.8)+
  theme_few()+
  labs(
       y = "Mass (g)",
       x = "Date")
```


#### April albedo
```{r}
albedo_april <- Energy_1hr_clean %>% 
  filter(month(date) == 04 & year(date) == 2015) %>% 
  select(date, albedo)

str(albedo_april)

albedo_april_clean <- dplyr::filter(albedo_april, albedo > 0 & albedo < 1)

albedo_plot <- albedo_april_clean %>% 
  filter(hour(date) > 13 & hour(date) < 24) %>%
  ggplot(aes(x = date, y = albedo))+
  geom_point() +
  geom_vline(xintercept = mdy_hms('04/01/2015 02:00:00'))+
  geom_vline(xintercept = mdy_hms('04/08/2015 02:00:00'))+
  geom_vline(xintercept = mdy_hms('04/14/2015 02:00:00'))

albedo_plot 


plotly::ggplotly(albedo_plot)

april_D2 <- albedo_april_clean %>% 
  filter(day(date) >= 6 & day(date) <= 13)
```

## Visualize Energy budget data
```{r, fig.cap= " Figure 1. Downdwelling long wave radiation measured at SASP for each year with dust on snow records. "}

Energy_clean %>% 
  ggplot(aes(x = Year, y = Pyrgeom_W, fill = Year))+
  geom_boxplot()+
  # Color Palette
  scale_fill_brewer(palette = "Spectral")+
  # Placing points on top of the box plot to see distribution
  #geom_jitter(size = 0.3)+
  # Titles
  labs(title ="Trends of Downdwelling Longwave Radiation",
       x = "Year",
       y = "Downward Longwave Raditiation (W/m2)")+
  # Break the graphs up into separate years
  facet_grid(~Year, scales = "free_x")+
  theme(
    text = element_text(size = 12),           # X axis labels
    legend.position = "none",                 # Remove the legend
    plot.title = element_text(hjust = 0.5)    # Move the title to the middle
  )
```



# Combining Data Sets --------
```{r}
view(Energy_1hr_clean)
view(dust_mass)

dust_bef_aft <- dust_mass %>% 
  mutate(DateClean = as.Date(DateClean, format = "%m/%d/%Y")) %>% 
  mutate(before = DateClean - 7)
```


